The Effects of weather on maize yield: New evidence from Kenya

### 2018-12-25 writing/draft3/draft18.tex
writing

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects are multiplicative. The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix. The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised. For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as positive effect. The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.

@@

The estimates of the model of the whole sample are summarised in the first two columns of table~\ref{MainEst}. Consistently with our assumptions, seasonal precipitation and average temperature have significant effects on yield. As we expected, the effect of temperature is negative. The linear term of precipitation is positive while the squared term of precipitation is negative and as we can see in figure~\ref{MarEff1}(a), the plot of the marginal effects of precipitation is hill-shaped. That is, the effect of precipitation is negative and increasing if the amount of precipitation is below its mean and the marginal effect increases with the amount of precipitation until it reaches $1.403$ times its standard deviation. At this point, the effect starts to decrease with increasing precipitation and it becomes negative again once the amount of precipitation reaches $2.805$ times its standard deviation. Therefore, too  much rain is harmful for agricultural yield as well as not enough rain. This is intuitive as extreme rainfall can cause flooding which is damaging for yield.

Besides cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and also length and number of dry spells were significant (see the first column of table~\ref{MainEst}). As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear. In accordance with our expectation, the yield decreases with increasing CV of precipitation. \hl{However, the result which we found for CV of temperature is a bit peculiar as the yield seem to increase with increasing variability of temperature. This is probably because...}

@@

We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}). As for the subsample of the ASAL counties, the only significant fixed effects included average seasonal temperature and length and number of dry spells, all of which were negative. Interestingly, different predictors were  found significant for the subsample of the non-ASAL counties. These included seasonal precipitation and its square (the first one was positive and the second one negative..\textcolor {red}{here}. )

>>>>

The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects are multiplicative. The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix. The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised. For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as a positive effect. If, for example, a marginal effect of an explanatory variable $x$ was $1.1$ when $x$ is equal to its one standard deviation, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).

The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.

@@

The estimates of the model of the whole sample are summarised in the first two columns of table~\ref{MainEst}. Consistently with our assumptions, seasonal precipitation and average temperature have significant effects on yield. As we expected, the effect of temperature is negative. The linear term of precipitation is positive while the squared term of precipitation is negative and as we can see in figure~\ref{MarEff1}(a), the plot of the marginal effects of precipitation is hill-shaped. That is, the effect of precipitation is negative and increasing if the amount of precipitation is below its mean and the marginal effect increases with the amount of precipitation until it reaches $1.403$ times its standard deviation. At this point, the effect starts to decrease with increasing precipitation and it becomes negative again once the amount of precipitation reaches $2.805$ times its standard deviation. Therefore, too  much rain is harmful for agricultural yield as well as not enough rain. This is intuitive mainly because extreme rainfall can cause flooding which is damaging for yield.

Other fixed effects which we found significant include CV of temperature and precipitation and length and number of dry spells (see the first column of table~\ref{MainEst}). As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear. In accordance with our expectation, the yield decreases with increasing CV of precipitation. \hl{However, the result which we found for CV of temperature is a bit puzzling as the yield seem to increase with increasing variability of temperature. This is probably because...}

@@

We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}). Interestingly, different predictors were  found significant for the two subsamples. Regarding the subsample of the ASAL counties, the only significant fixed effects are average seasonal temperature and length and number of dry spells. All three variables have a negative sign. On the other hand, the variables which were found significant for the subsample of the non-ASAL counties include seasonal precipitation and its square with a positive and a negative sign, respectively and CV of temperature which has a positive sign as in the case of the entire sample. According to the F-statistic, the most important explanatory variables for the ASAL counties are number of dry spells with F-statistic equal to $8.032$ and length of the longest dry spell with F-value equal to $6.841$. The F-statistic of the average temperature is $5.409$ and it lower than zero for the remaining predictors. However, according to the magnitude of the standardised coefficient, the average temperature would be the most important as the coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively. With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its square. The F-statistic of the squared values of precipitation is $23.384$ and its value for the linear term of precipitation is $19.140$, while the variable with the third highest F-statistic is CV of temperature with the F-statistic equal to $5.372$, hence, the difference in importance of the precipitation and CV of temperature is relatively large. The standardised coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised coefficient of CV of temperature is $0.058$. In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of the coefficient is $-0.128$. However, average temperature is insignificant according to the Satterthwaite's p-value and its F-statistic is also relatively small, in particular $1.528$.

To sum up, in the ASAL counties temperature is much more important than in the non-ASAL counties where temperature is not significant. On the other hand, precipitation is not significant in the ASAL counties while it is strongly significant in the non-ASAL counties. This is illustrated by figure~\ref{MarEff1}. The plot of the marginal effects of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(b)) while is steep and hill-shaped for the non-ASAL counties similarly as in the case of the whole sample. As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f), the marginal effects of temperature are linear and decreasing for the case of ASAL and non-ASAL counties. Marginal effects of the remaining predictors of the models for the two subsamples are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
